---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# The California Department of Education in R

The California Department of Education provides ample data. Now, that data is available in R.

## Installation

```{r, eval=FALSE}
devtools::install_github("daranzolin/rCAEDDATA")
library(rCAEDDATA)
```


## Available Datasets

* [Cohort Outcome Data ("cohorts")](http://www.cde.ca.gov/ds/sd/sd/filescohort.asp) -- California Longitudinal Pupil Achievement Data System (CALPADS) cohort outcome data reported by race/ethnicity, program participation, and gender.

* [Dropouts by Race and Gender ("dropouts")](http://www.cde.ca.gov/ds/sd/sd/filesdropouts.asp) -- Data for grade seven through twelve dropouts and enrollment by race/ethnic designation and gender by school.

* [English Learners by Grade and Language ("english_learners")](http://www.cde.ca.gov/ds/sd/sd/fileselsch.asp) -- Data for English learners (ELs) by grade, language, and school.

* [Enrollment by School ("enrollments")](http://www.cde.ca.gov/ds/sd/sd/filesenr.asp) -- Data for school-level enrollment by racial/ethnic designation, gender, and grade.

* [Student Poverty FRPM ("frpm")](http://www.cde.ca.gov/ds/sd/sd/filessp.asp) -- Data for students eligible for Free or Reduced Price Meals (FRPM).

* [Graduates by Race and Gender ("graduates")](http://www.cde.ca.gov/ds/sd/sd/filesgrads.asp) -- Data for graduates and graduates meeting University of California (UC)/California State University (CSU) entrance requirements by race/ethnic designation and gender by school.

* [Primary and Short-Term Enrollment ("primary_and_short_term")](http://www.cde.ca.gov/ds/sd/sd/filesenrps.asp) -- Data for primary and short-term school-level enrollment by racial/ethnic designation, gender, and grade.

## Examples

```{r, fig.width=12}
library(rCAEDDATA)
library(tidyverse)
data("graduates")
graduates %>% 
  group_by(YEAR) %>% 
  summarize(total_grads = sum(GRADS),
            UC_CSU = sum(UC_GRADS),
            non_UC_CSU = total_grads - UC_CSU) %>%
  select(-total_grads) %>% 
  gather(Eligibility, Graduates, -YEAR) %>% 
  ggplot(aes(YEAR, Graduates, fill = Eligibility)) +
  geom_bar(stat = "identity", color = "black") +
  labs(x = "Year",
       y = "Graduates",
       title = "California High School Graduates, 1992-2016") +
  scale_y_continuous(labels = scales::comma) +
  scale_fill_manual(values = c("yellow", "lightblue")) +
  theme_minimal() 
  
```

